---
title: "Bioinformatics Lecture 11"
author: "Cassandra Chow"
date: "May 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##PDB Statistics
Download CSV file from PDB database

```{r}
pdb.stats <- read.csv("Data Export Summary.csv")
```

Calculating percentages of the amount of data in the PDB database
```{r}
percent <- (pdb.stats$Total/ sum(pdb.stats$Total))*100
names(percent) <- pdb.stats$Experimental.Method
percent
```
X-Ray takes about 89.5% of the entries on PDB

Caclulating the percentage of proteins 
```{r}
percent.p <- (pdb.stats$Proteins/sum(pdb.stats$Proteins))*100
names(percent.p) <- pdb.stats$Experimental.Method
percent.p
```
90% of X-ray readings are protein structures

There are currently 1103 HIV-1 Structures on the PDB website

##Accessing 1HSG PDB File
Installing files
```{r}
library(bio3d)
```

List the functions of this package 
```{r}
pdb <- read.pdb("1hsg")
```

Calling out C-alpha only file
```{r}
ca.inds <- atom.select(pdb, "calpha")
```

Checking if we have what we want 
```{r}
head(pdb$atom[ca.inds$atom,])
```

Take out our indices and trim our PDB
```{r}
ca.pdb <- trim.pdb(pdb, inds = ca.inds)
```

Write this out 
```{r}
write.pdb(ca.pdb, file = "c_alpha_only.pdb")
```