---
title: "Hands On Analysis of Cancer Cells"
author: "Cassandra Chow"
date: "May 1, 2018"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Preparing the Data 
We will be importing data into R
```{r}
url <- "https://bioboot.github.io/bimm143_W18/class-material/WisconsinCancer.csv"
wisc.df <- read.csv(url)
```

Now, we have to convert this into a matrix
```{r}
wisc.data  <- as.matrix(wisc.df[, 3:ncol(wisc.df)])
rownames(wisc.data) <- wisc.df$id
```

Checking up diagnostics
```{r}
diagnostics <- as.numeric(wisc.df$diagnosis == "M")
```


##Exploring Data Analysis 
Q1. How many observations are in this dataset?
```{r}
nrow(wisc.data)
```

Q2. How many variables/features in the data are suffixed with _mean?
```{r}
no_of_variables <- max(grep("mean", colnames(wisc.data)))
paste("The number of variables with the suffixed mean is", no_of_variables, sept = " ")
```


Q3. How many of the observations have a malignant diagnosis?
```{r}
malignant <- sum(diagnostics)
paste("The number of observations of those that have a malignant diagnosis is",
      malignant, sep = " ")
```

##Performing PCA
We will be checking the column means and standard deviations
```{r}
colMeans(wisc.data)
apply(wisc.data, 2, sd)
```
As you can tell, this is not pretty looking

Hence, we are trying this prcomp function. 
Note that we have to omit the last column due to nonnumeric values.
```{r}
wisc.pr <- prcomp(wisc.data[,1:30], scale. = TRUE)
```

Checking the summary of wisc.pr
```{r}
summary(wisc.pr)
```

Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
The answer is 0.4427

Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
We need just three componenets tha describe at least 70%.

Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
We need seven component that describe at least 90%. 

##Interpreting PCA Results
We will use the biplot function
```{r}
biplot(wisc.pr)
```

Using the scatter plot, add 1 to diagnostics since the value of zero will be white
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = (diagnostics + 1), 
     pch = as.vector(wisc.df$diagnosis), xlab = "PC1", ylab = "PC2")
```

Repeat for components 1 and 3
```{r}
plot(wisc.pr$x[, c(1, 3)], col = (diagnostics + 1), pch = as.vector(wisc.df$diagnosis),
     xlab = "PC1", ylab = "PC3")
```
We see that this plot is more scattered than the previous one. 

Alternative scree plot of the same data, note data driven y-axis
```{r}
pr.var <- (wisc.pr$sdev)^2 #this will be used later

pve <- pr.var/sum(pr.var)
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

Plot cumulative proportion of variance explained
```{r}
plot(pr.var, xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

##Selecting number of clusters
We need to scale the data and finds the clusters
```{r}
data.scaled <- scale(wisc.data[,-31])

data.dist <- dist(data.scaled)

wisc.hclust <- hclust(data.dist)
```

Working to make this 
```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, 4)
table(wisc.hclust.clusters, diagnostics)
```

##K-means clustering and comparing results
Form the k-means claster 
```{r}
wisc.km <- kmeans(scale(wisc.data[,-31]), centers = 2, nstart = 20)
```

Next, we need to create a table
```{r}
table(wisc.km$cluster, diagnostics)
```

We also want to compare k-means clustering and hclust
```{r}
table(wisc.hclust.clusters, wisc.km$cluster)
```

##Clustering on PCA results
Use the distance along the first 7 PCs for clustering i.e. wisc.pr$x[, 1:7]
```{r}
#finding the distance
wisc.pr.dist <- dist(wisc.pr$x[, 1:7])

#using the hclust function
wisc.pr.hclust <- hclust(wisc.pr.dist, method = "complete")
```

Comparing it now using k-means clust
```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=4)
```

Now, let's look at the results
```{r}
table(wisc.pr.hclust.clusters, diagnostics)
```


